Texture Mapping 3D Models of Indoor Environments with Noisy Camera Poses

### 2012-06-29 3dimpvt/paper.tex
references in, too long

%%%%%%%%% ABSTRACT
\begin{abstract}
  Automated 3D modeling of building interiors is useful in
  applications such as virtual reality and environment
  mapping. Applying textures to these models is an important step in
  generating photorealistic visualizations of data collected by
  modeling systems.  Camera pose recovery in such systems often
  suffers from inaccuracies, resulting in visible discontinuties when
  different images are projected adjacently onto a plane for
  texturing. We propose two approaches for reducing discontinuities in
  texture mapping 3d models made of planar surfaces. The first one is
  tile based and can be used where camera axes are at arbitrary angles
  with respect to the plane normals of the surfaces. The second one
  results in a more seamless texture but is only applicable where
  camera axes for imagery are closely aligned with plane normals. The
  effectiveness of our approaches are demonstrated on two indoor
  datasets.
\end{abstract}

@@

Mounting such devices on a human-carried platform provides unique
advantages over vehicular-based systems on wheels in terms of agility
and portability, but can also result in much larger localization
error. As a result, common methods for texture mapping generally
produce poor results.

In this paper, we present a number of approaches to texture mapping 3D
models of indoor environments made of planar surfaces in the presence
of uncertainty and noise in camera poses. In particular, we consider a
human-operated backpack system with a number of laser range scanners
as well as 2 cameras facing left and right, each equipped with fisheye
lenses reaching an approximately 180$^{\circ}$ field of view. With the
laser scanners active, the human operator wearing the backpack takes
great care to walk a path such that every wall in the desired indoor
environment is traversed and scanned lengthwise at least once.

Using data collected by the onboard sensors and applying multiple
localization and loop-closure algorithms \cite{chen2010indoor,
  kua2012loopclosure, liu2010indoor}, the backpack is then localized
over its data collection period. This requires recovering the 6
degrees of freedom for the backpack as well as all its sensors,
including the cameras mounted on it\footnote{In this paper, we use the
  terms localization and pose recovery interchangeably, in that they
  both refer to recovering position and orientation.}. Once this is
complete, the data from the laser range scanners is used to generate a
3D point cloud of the surrounding environment.  Approximate normal
vectors for each point in the point cloud are then calculated by
gathering neighboring points within a small radius and processing them
through principal component analysis. These normal vectors allow for
the classification and grouping of adjacent points into structures
such as walls, ceilings, floors, and staircases. A RANSAC algorithm is
then employed to fit polygonal planes to these structured groupings of
points, resulting in a fully planar model
\cite{sanchez2012point}. This model, consisting of multiple 2D
polygonal planes in 3D space, along with the set of images captured by
the backpack's cameras and their noisy 3D poses, can be considered the
input to our texture mapping problem.

Our problem is different from others in that we are more interested in
creating accurate and seamless textures than in recovering geometry
out of our photos, as is a main concern in many related works
\cite{bernardinimultiplescans, debevechybrid,
  liu2010indoor}. Additionally, our surfaces to be textured are planes
captured from close distances, so we do not always have many visual
reference points to work with, as is the case when texturing detailed objects
at high resolutions \cite{bernardinimultiplescans, wangmultipleviews}, or when creating
large scene panoramas \cite{agarwalapanoramas, bernardinimultiplescans,
  debevecviewdependent}. Instead, we need to adjust image projections
relative to each other, while still taking the noisy camera
poses provided by our localization algorithms into account. Using these camera
poses, we also aim for complete automation and improved runtime over
non-localized processes, such as structure from motion or image-based
rendering \cite{agarwalapanoramas, debevechybrid,
  snavelyphototourism}.

The remainder of the paper is organized as follows. Section
\ref{sec:simpleTextureMapping} describes the general problem of
texture mapping and examines two variants of a tile-based mapping
approach and their shortcomings.  Section \ref{sec:existingApproaches}
demonstrates two previous attempts at improving texture alignment, and
demonstrates their inadequacies for our datasets. Section
\ref{sec:proposedApproach} presents our proposed approach to texture
mapping, combining an improved localization refinement process with
two image selection approaches. Section
\ref{sec:resultsAndConclusions} contains results and conclusions.

@@

The geometry of the texture mapping process for a plane is shown in
Figure \ref{fig:projection}.  As described earlier, we are provided
with a set of $M$ images with which we must texture our target
plane. Each image has a camera matrix $P_i$ for $i=1..M$, which
translates a 3D point in the world coordinate system to a 2D point or
pixel in image $i$'s coordinates. A camera matrix $P_i$ is composed of
the camera's intrinsic parameters, such as focal length and image
center, as well as extrinsic parameters which specify the rotation and
translation of the camera's position in 3D world coordinates at the
time that image $i$ is taken. These extrinsic parameters are
determined by the backpack hardware and localization algorithms
\cite{chen2010indoor, liu2010indoor, kua2012loopclosure} and are quite
noisy.

Our goal is to texture each plane using images captured by the
backpack system, while eliminating any visual discontinuities or seams
that would suggest that the plane's texture is not composed of a
single continuous image.

@@

Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
we can texture the plane by discretizing it into small square tiles,
generally about 5 pixels across, and choosing an image to texture each
tile with. We choose to work with rectangular units to ensure that
borders between any two distinct images in our final texture are
either horizontal or vertical. Since most environmental features
inside buildings are horizontal or vertical, any seams in our texture
intersect them minimally and are likely to be less noticeable.

@@

Once we have a list of candidate images for $t$, we must define a
scoring function in order to compare images and objectively select the
best one. Since camera localization errors compound over distance, we
wish to minimize the distance between cameras and the plane they
texture. Additionally, we desire images that are projected
perpendicularly onto the plane, maximizing the resolution and amount
of useful texture available in their projections. In other words, we
wish to minimize the angle between the plane normal and the camera
axis for images selected for texture mapping. These two criteria can
be met by maximizing the function $\frac{1}{d} (-1 \cdot \vec{C})
\cdot \vec{N}$ with the parameters shown in Figure
\ref{fig:scoringFunction}. Specifically, $d$ is the distance between
the centers of a camera and a tile, and $\vec{N}$ and $\vec{C}$ are
the directions of the plane's normal and the camera axis respectively.

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{scoringFunction.pdf}
  \caption{We minimize camera angle $\alpha$ and distance $d$ by maximizing the scoring function $\frac{1}{d} (-1 \cdot \vec{C})
\cdot \vec{N}$}
  \label{fig:scoringFunction}
\end{figure}

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=2.3in]{wall2_naive.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=2.3in]{wall2_cache.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=2.3in]{wall2_cache_shift.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=2.3in]{wall2_dynprog_noblend.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=2.3in]{wall2_cache_shift_blend.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=2.3in]{wall2_dynprog.jpg}}
  \caption{(a): Direct mapping. (b): Mapping with caching. (c):
    Mapping with caching after image alignment. (d): Seam minimization after image alignment). (e): (c) after blending. (f): (d) after blending.}
  \label{fig:compareAll}
\end{figure}

@@

\subsection{Mapping with Caching}
\label{sec:mappingWithCaching}
Since discontinuities occur where adjacent tiles select non-aligned
images, it makes sense to take into account image selections made by
neighboring tiles while selecting the best image for a given tile. By
using the same image across tile boundaries, we can eliminate a
discontinuity altogether. If this is not possible because a tile is
not visible in images chosen by its neighbors, using similar images is
likely to result in less noticeable discontinuities.

Similar to a caching mechanism, we select the best image for a tile
$t$ by searching through two subsets of images for a good candidate,
before searching through the entire set. The first subset of images is
those selected by adjacent tiles that have already been textured. We
must first check which of these images can map to $t$, and then of
those, we make a choice according to the scoring function in Figure
\ref{fig:scoringFunction}. Before reusing this image, we ensure it
meets a threshold, which we set unconditionally to $\alpha <
45^\circ$, to be considered a viable image, with $\alpha$ as the
camera angle as shown in Figure \ref{fig:scoringFunction}.

If no satisfactory image is found in the first subset, we then check
our second subset of images, which consists of images that were taken
near the images in the first subset, both spatially and
temporally. These images are not the same as the ones used for
neighboring tiles, but they were taken at a similar location and time,
suggesting that their localization and projection are very
similar. Again, if no good image is found according to the same
threshold, we must then search the entire set of candidate images,
selecting based on the same scoring function from Figure
\ref{fig:scoringFunction}.

The results of this caching approach are shown in Figure
\ref{fig:compareAll}(b). As compared to Figure
\ref{fig:compareAll}(a), discontinuities are reduced overall, but the
amount of remaining seams suggests that image selection alone cannot
produce seamless textures. Camera matrices, or the image projections
themselves have to be adjusted in order to reliably generate clean
textures.

\section{Existing Approaches to Image-Aligned Texture Mapping}
\label{sec:existingApproaches}
In order to produce seamless texture mapping, either camera matrices
need to be refined such that their localization is pixel accurate, or
image stitching techniques need to be applied to provide this
illusion.

Before examining these approaches, we first obtain a set of images to
work with.  Rather than perform camera or image adjustments across the
many thousands of images acquired in a typical data collection, we opt
to work with the more limited set of images corresponding to those
chosen by the direct mapping approach, without caching (Section
\ref{sec:directMapping}. This set of images constitutes a good
candidate set for generating a final seamless texture since it meets
three important criteria. First, each tile on our plane is covered by
at least one image in this set; this ensures no holes in our final
texture. Second, images are all selected according to the scoring
function in Figure \ref{fig:scoringFunction}, ensuring good camera
distances and angles. Third, as a side result of the scoring function,
selected images are only viable candidates for the tiles near their
center of projection. Thus, there should be plenty of overlap between
selected images, allowing for some degree of shifting without
resulting in holes, as well as area for blending between them. With
this set of images, we now review two existing approaches towards
refining and combining their projections.

\subsection{Image Mosaicing}
\label{sec:imageMosaicing}

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{panoMy.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=3in]{Graph_crop.pdf}}
  \caption{Both image mosaicing (a) and the graph-based localization refinement algorithm from
    \cite{chen2010indoor} (b) suffer from the problem of compounding
    errors.}
  \label{fig:mosaic3D}
\end{figure}

When images of a plane are taken from arbitrary overlapping positions,
they are related by homography \cite{hz}. Thus, existing
homography-based image mosaicing algorithms are applicable
\cite{brown2007automatic}. However, errors can compound when long
chains of images are mosaiced together using these approaches. For
example, a pixel in the $n$th image in the chain must be translated
into the first image's coordinates by multiplying by the $3\times3$
matrix $H_1 H_2 H_3 ... H_n$. Any error in one of these homography
matrices is propagated to all further images until the chain is
broken. For some chains of images this can happen almost immediately
due to erroneous correspondence matches and the resulting image mosaic
is grossly misshapen.

Figure \ref{fig:mosaic3D}(a) shows the output of the AutoStitch software
package which performs homography-based image mosaicing. This plane is
nearly a best-case scenerio with many features spread uniformly across
it. Even so, the mosaicing produces errors that causes straight lines
to appear as waves on the plane, despite the fact that it was
generated after careful hand tuning. Many planes with fewer features
simply failed outright. Thus, image mosaicing is not a robust enough
solution for reliably texture mapping our dataset.

\subsection{Image-Based 3D Localization Refinement}
\label{sec:imageBased3DRefinement}

Another approach is to refine the camera matrices using image
correspondences to guide the process. Each image's camera matrix has 6
degrees of freedom that can be adjusted. Previous work on this problem
attempted to refine camera matrices by solving a non-linear
optimization problem \cite{liu2010indoor}. This process is specific to
the backpack system which generated our dataset, as it must be run
during backpack localization\cite{liu2010indoor,
  chen2010indoor}. This approach suffers from a similar
error propagation problem as shown in Figure \ref{fig:mosaic3D}(b).

@@

\section{Proposed Method for Seamless Texture Mapping}
\label{sec:proposedApproach}
In this section, we will describe our proposed method for seamless
texture mapping. Our approach consists of 4 steps, as seen in Figure
\ref{fig:pipeline}.  First, we choose a subset of $M$ images to work
with via the direct mapping approach of Section
\ref{sec:directMapping}. Next, we project these images onto the plane
and perform rotation and shifting in order to maximize SIFT feature
matches in overlapping areas between these projections. We then select
which textures to be used either with the tile caching method from
Section \ref{sec:mappingWithCaching}, or with the more specialized
method to be described in Section \ref{sec:seamMinimization}. This
choice depends on the availability of head-on images for a given
plane. We then finish by applying linear alpha blending to smooth out
any remaining seams.

\subsection{Image Projection and Rotation}
\label{sec:projectionAndRotation}
We begin with the projection of all images onto the plane. This is
done in the same way as the approaches in Section
\ref{sec:simpleTextureMapping}. We then perform rotations on these
projections, as adjacent images with different orientations result in
strong discontinuities.

These rotations are accomplished by using Hough transforms, which
detect the presence and orientation of linear features in our
images. Rather than match the orientation of such features in each
image, we simply apply rotations such that the strongest near-vertical
features are made completely vertical. This is effective for vertical
planes in indoor models, since they usually consist of parallel
vertical lines corresponding to doors, wall panels, rectangular
frames, etc. If features in the environment are not vertical, or are
not parallel to eachother, this step is skipped.

\subsection{Image Pose Refinement}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align overlapping images by searching for
corresponding points between all pairs of overlapping images using
SIFT feature matching \cite{lowe1999object}. SIFT matches
determine $dx$ and $dy$ distances between each pair of features for
two images on the plane, though these distances may not always be the
same for different features.

Since indoor environments often contain repetitive features such as
floor tiles or doors, we need to ensure that our SIFT-based distances
are reliable. In order to mitigate the effect of incorrect matches and
outliers, the RANSAC framework \cite{fischler1981random} is used for a
robust estimate of the optimal $dx_{i,j}$ and $dy_{i,j}$ distances
between two images $i$ and $j$. The RANSAC framework handles the
consensus-building machinery, and requires a fitting function and a
distance function. For this application, the fitting function simply
finds the average distance between matches in a pair of images. The
distance function for a pair of points is chosen to be the difference
between those points' SIFT match distance and the average distance
computed by the fitting function. We use a 10 pixel outlier
threshold, so that SIFT matches are labeled as outliers if
their horizontal or vertical distances are not within 10 pixels of the
average distance computed by the fitting function.

We now use the $dx_{i,j}$ and $dy_{i,j}$ distances between each pair
of images to refine their positions using least squares. Recall that
there are a total of $M^{2}$ possible pairs of images, though we only
generate distances between images that overlap at SIFT feature
points. Given these distances and the original image location
estimates, we can solve a least squares problem
($\textrm{min}_{\vec{\beta}} ||A \vec{\beta} - \vec{\gamma}||_2^2 $)
to estimate the location of the images on the plane. The
$M$-dimensional vector $\vec{\beta}$ represents the unknown $x$
location of each image on the plane for $1 \dots M$. The optimal $x$
and $y$ locations are obtained in the same way, so we only consider
the $x$ locations here:

@@

The $\vec{\beta}$ that minimizes $||A \vec{\beta} -
\vec{\gamma}||_2^2$ results in a set of image locations on the plane
that best honors all the SIFT-based distance measurements between
images. In practice there are often cases where there is a break in
the chain of images, meaning that no SIFT matches were found between
one segment of the plane and another. In this case we add rows to the
$A$ matrix and observations to the $\vec{\gamma}$ vector that contain
the original noisy $x$ and $y$ distance estimates generated by the
localization algorithm \cite{chen2010indoor, liu2010indoor}. Another
way to do this is to add rows for all neighboring pairs of images and
solve a weighted least squares problem where the SIFT distances are
given a higher weight i.e. 1, and the noisy distances generated by the
localization algorithm \cite{chen2010indoor, liu2010indoor} are given
a smaller weight i.e. 0.01.

After completing this same process for the $y$ dimension as well, and
making the resultant shifts, our images overlap and match each other
with far greater accuracy. Applying the tile caching method from
Section \ref{sec:mappingWithCaching} on these re-localized images
results in the significant improvements shown in Figure
\ref{fig:compareAll}(c), as compared to Figure
\ref{fig:compareAll}(b).

@@

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=1in]{projectionWall.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=1.1in]{projectionCeiling.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=0.9in]{projectionWallCrop.pdf}}
  \caption{(a): Undistorted fisheye images for vertical planes are
    tilted, but their effective camera axis is more or less normal to
    the plane. (b): The effective camera axis for ceilings is at an
    angle with respect to the plane normal. (c): Wall images are
    cropped to be rectangular.}
  \label{fig:projectionAngles}
\end{figure}

As mentioned earlier, the mobile backpack system used to capture data
in this paper contains 2 cameras with $180^\circ$ fisheye lenses
facing to the right and left. Since most human operators do not carry
the backpack in a perfectly upright position and are bent forwards at
15 to 20 degrees with respect to the vertical direction, the
undistorted fisheye images are head on with respect to vertical walls,
but at an angle with respect to horizontal floors and ceilings. This
is depicted in Figure \ref{fig:projectionAngles}.

>>>>

%%%%%%%%% ABSTRACT
\begin{abstract}
  Automated 3D modeling of building interiors is useful in
  applications such as virtual reality and environment
  mapping. Applying textures to these models is an important step in
  generating photorealistic visualizations of data collected by
  modeling systems. Camera pose recovery in such systems often suffers
  from inaccuracies, resulting in visible discontinuties when images
  are projected adjacently onto a plane for texturing. We propose two
  approaches for reducing discontinuities in texture mapping 3D models
  made of planar surfaces. The first one is tile based and can be used
  for images and planes at arbitrary angles relative to each
  other. The second one, which we refer to as our seam minimization
  approach, results in a more seamless texture but is only applicable
  where camera axes for images are closely aligned with plane normals
  of the surfaces to be textured. The effectiveness of our approaches
  are demonstrated on two indoor datasets.
\end{abstract}

@@

Mounting such devices on a platform carried by an ambulatory human
provides unique advantages over vehicular-based systems on wheels in
terms of agility and portability, but can also result in larger
localization error \cite{liu2010indoor}. As a result, common methods
for texture mapping generally produce poor results.

In this paper, we present a number of approaches to texture mapping 3D
models of indoor environments made of planar surfaces in the presence
of uncertainty and noise in camera poses. In particular, we consider a
human-operated backpack system with a number of laser range scanners
as well as 2 cameras facing left and right, each equipped with fisheye
lenses reaching an approximately 180$^{\circ}$ field of view and
taking photos at a rate of 5 Hz. With the cameras and laser scanners
active, the human operator wearing the backpack takes great care to
walk a path such that every wall in the desired indoor environment is
traversed and scanned lengthwise at least once, resulting in many
thousands of captured images.

Applying multiple localization and loop-closure algorithms on the data
collected by the ondoor sensors \cite{chen2010indoor,
  kua2012loopclosure, liu2010indoor}, the backpack is
localized\footnote{In this paper, we use the terms localization and
  pose recovery interchangeably, in that they both refer to recovering
  position and orientation.}  over its data collection period. This
requires recovering the 6 degrees of freedom for the backpack as well
as all its sensors, including the cameras rigidly mounted on it. Once
this is complete, the data from the laser range scanners is used to
generate a 3D point cloud of the surrounding environment and a 3D
planar model is generated \cite{sanchez2012point}. To do so, we
approximate normal vectors for each point in the point cloud by
gathering neighboring points within a small radius and processing them
through principal component analysis. These normal vectors allow for
the classification and grouping of adjacent points into structures
such as walls, ceilings, floors, and staircases. A RANSAC algorithm is
then employed to fit polygonal planes to these structured groupings of
points, resulting in a fully planar model. This model, consisting of
multiple 2D polygonal planes in 3D space, along with the set of images
captured by the backpack's cameras and their noisy 3D poses, can be
considered the input to our texture mapping problem.

This problem straddles the fields of image-based localization
refinement as well as image alignment and mosaicing, and thus our
proposed solution shares many similarities with past research in those
areas, which will be examined in Section \ref{sec:existingApproaches}. However, given our knowledge of approximate camera poses as well as the 3D geometry of the environment to be textured, we can more accurately and efficiently texture our planes.

The remainder of the paper is organized as follows. Section
\ref{sec:simpleTextureMapping} describes two variants of a tile-based
mapping approach and their shortcomings.  Section
\ref{sec:existingApproaches} discusses past research in related
fields, demonstrates two previous attempts at improving texture
alignment, and exhibits their inadequacies for our datasets. Section
\ref{sec:proposedApproach} presents our proposed seam minimization
approach to texture mapping. Section \ref{sec:resultsAndConclusions}
contains results and conclusions.

@@

The geometry of the texture mapping process for a plane is shown in
Figure \ref{fig:projection}.  As described earlier, we are provided
with a set of $M$ images with which we must texture the target
plane. Each image has a camera matrix $P_i$ for $i=1..M$, which
translates a 3D point in the world coordinate system to a 2D point or
pixel in image $i$'s coordinates. A camera matrix $P_i$ is composed of
the camera's intrinsic parameters, such as focal length and image
center, as well as extrinsic parameters which specify the rotation and
translation of the camera's position in 3D world coordinates at the
time that image $i$ is taken. These extrinsic parameters are
determined by the backpack hardware and localization algorithms
\cite{chen2010indoor, liu2010indoor, kua2012loopclosure} and are quite
noisy.

Because the backpack system takes photos at a rate of 5 Hz, thousands
of images are available for texturing each plane. Our goal in creating
a texture mapping process is to decide which of these images should be
used, and where their contents should map onto the plane, in order to
eliminate any visual discontinuities or seams that would suggest that
the plane's final texture is not composed of a single continuous image.

@@

Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate, we can texture the plane by discretizing it into small square tiles, generally about 5 pixels across, and choosing an image to texture each tile.

We choose to work with rectangular units to ensure that borders
between any two distinct images in our final texture are either
horizontal or vertical. Since most environmental features inside
buildings are horizontal or vertical, any seams in our texture
intersect them minimally and are likely to be less noticeable.

@@

Once we have a list of candidate images for $t$, we define a scoring
function in order to objectively select the best image. Since camera
pose errors compound over distance, we wish to minimize the distance
between cameras and the plane they texture. Additionally, we desire
images that are projected perpendicularly onto the plane, maximizing
the resolution and amount of useful texture available in their
projections, as well as minimizing any parallax effects. In other words, we wish to minimize the angle between the
plane normal and the camera axis for images selected for texture
mapping. These two criteria can be met by maximizing the function
$\frac{1}{d} (-1 \cdot \vec{c}) \cdot \vec{n}$ as shown in Figure
\ref{fig:scoringFunction}. Specifically, $d$ is the distance between
the centers of a camera and a tile, and $\vec{n}$ and $\vec{c}$ are
the directions of the plane's normal and the camera axis respectively.

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{scoringFunction.jpg}
  \caption{We minimize camera angle $\alpha$ and distance $d$ by
    maximizing the scoring function $\frac{1}{d} (-1 \cdot \vec{c})
    \cdot \vec{n}$}
  \label{fig:scoringFunction}
\end{figure}

\begin{figure}[h!]
  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.9in]{wall2_naive.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.9in]{wall2_cache.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.9in]{wall2_cache_shift.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.9in]{wall2_dynprog_noblend.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.9in]{wall2_cache_shift_blend.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.9in]{wall2_dynprog.jpg}}
  \caption{(a) Direct mapping. (b) Mapping with caching. (c) Mapping
    with caching after image alignment. (d) Seam minimization after
    image alignment). (e) same as (c) with blending. (f) same as (d)
    with blending.}
  \label{fig:compareAll}
\end{figure}

@@

\subsection{Mapping with Caching}
\label{sec:mappingWithCaching}
Since discontinuities occur where adjacent tiles select non-aligned
images, it makes sense to take into account image selections made by
neighboring tiles while texture mapping a given tile. By using the
same image across tile boundaries, we can eliminate a discontinuity
altogether. If this is not possible because a tile is not visible in
images chosen by its neighbors, using similar images across a tile boundary will result in minimal discontinuities.

Similar to a caching mechanism, we select the best image for a tile
$t$ by searching through two subsets of images for a viable candidate,
before searching through the entire set. The first subset of images is
those selected by adjacent tiles that have already been textured. We
must first check which of these images can map to $t$, and then of
those, we make a choice according to the scoring function in Figure
\ref{fig:scoringFunction}. Before reusing this image, we ensure it
meets the criteria $\alpha < 45^\circ$, in order to be considered a
viable image, with $\alpha$ as the camera angle as shown in Figure
\ref{fig:scoringFunction}.

If no satisfactory image is found in the first subset, we check the
second subset of images, consisting of those taken near the ones in
the first subset, both spatially and temporally. These images are not
the same as the ones used for neighboring tiles, but were taken at a
similar location and time, suggesting that their localization and
projection are very similar. Again, if no viable image is found
according to the same criteria, we search the entire set of candidate
images, selecting based on the same scoring function from Figure
\ref{fig:scoringFunction}.

The result of this caching approach is shown in Figure
\ref{fig:compareAll}(b). As compared to Figure
\ref{fig:compareAll}(a), discontinuities are reduced overall, but the
amount of remaining seams suggests that image selection alone cannot
produce seamless textures. Camera matrices, or the image projections
themselves have to be adjusted in order to reliably generate clean
textures.

\section{Existing Approaches to Image Alignment}
\label{sec:existingApproaches}
In order to produce seamless texture mapping, either camera matrices
need to be refined such that their localization is pixel accurate, or
image stitching techniques need to be applied to provide this
illusion.

Before examining these approaches, we first obtain a set of images to
work with.  Rather than perform camera or image adjustments across the
many thousands of images acquired in a typical data collection, we opt
to work with the more limited set of images corresponding to those
chosen by the direct mapping approach, without caching, of Section
\ref{sec:directMapping}. This set of images constitutes a reasonable
candidate set for generating a final seamless texture since it meets
three important criteria. First, each tile on the plane is covered by
at least one image in this set; this ensures no holes in the final
texture. Second, images are all selected according to the scoring
function in Figure \ref{fig:scoringFunction}, ensuring reasonable
camera distances and angles. Third, as a side result of the scoring
function, selected images are only viable for the tiles near their
center of projection. Thus, there should be plenty of overlap between
selected images, allowing for some degree of shifting without
resulting in holes, as well as area for blending between them. With
this set of images, we now review approaches towards refining and
combining their projections.

One way to align image projections is to refine their noisy camera
poses, such that they become pixel accurate. This can be done by using
image correspondences to adjust each image's camera matrix over 6
degrees of freedom (DOF) \cite{coorg1997matching,
  liu2010indoor, shum2000systems, szeliski1997creating}. Using our same modeling datasets, recent work
followed this approach by solving a non-linear optimization problem to
refine camera poses during backpack localization \cite{liu2010indoor,
  chen2010indoor}. This approach however suffers from error
accumulation and drift as shown in Figure \ref{fig:mosaic3D}(a), as
well as high complexity and runtime.

Refining camera poses is analogous to image alignment or mosaicing,
which perform translations on images in order to align them directly.
One common approach to image alignment is to iteratively match images
on intensity differences in selected patches, often in a spatial
hierarchy \cite{hager1998efficient, lucas1981iterative,
  shum2000systems, szeliski2006image}. Another widely-used approach is
to perform feature matching, which requires the detection of visual
features and their existence in multiple images
\cite{brown2007automatic, cho2003automatic,
  mikolajczyk2005performance, szeliski2006image,
  yun2003comprehensive}. Notably, both approaches rely on the presence
of multiple visual references, as when texturing detailed objects at
high resolutions \cite{bernardinimultiplescans, wangmultipleviews}, or
when creating large scene panoramas \cite{agarwalapanoramas,
  bernardinimultiplescans, debevecviewdependent,
  snavelyphototourism}. In contrast, our indoor datasets have a high
prevalence of bare walls and ceilings, resulting in few reference
points that can be used to judge alignment.

Additionally, our datasets often contain long chains of images, which
leads to error accumulation when image correspondences are not
accurate. For example, when matching a long chain of images through
homography, a pixel in the $nth$ image must be translated into the
first image's coordinates by multiplying by the $3\times3$ matrix $H_1
H_2 H_3 ... H_n$. Any error in one of these homography matrices is
propagated to all further images, resulting in drift. Figure
\ref{fig:mosaic3D}(b) shows the output of the AutoStitch software
package, which performs homography-based image mosaicing
\cite{autostitch}. Even with many features spread across this plane,
the mosaicing produces errors that causes straight lines to appear as
waves on the plane, despite the fact that it was generated after
careful hand tuning. Many planes with fewer features simply failed
outright.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{Graph_crop.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3in]{panoMy.jpg}}
  \caption{Texture alignment via (a) the graph-based localization
    refinement algorithm from \cite{chen2010indoor} and (b) image
    mosaicing}
  \label{fig:mosaic3D}
\end{figure}

\section{Seamless Texture Mapping}
\label{sec:proposedApproach}

@@

In this section, we describe our proposed method for seamless texture
mapping. Our approach consists of 4 steps, as seen in Figure
\ref{fig:pipeline}.  First, we choose a subset of $M$ images to work
with via the direct mapping approach of Section
\ref{sec:directMapping}. Next, we project these images onto the plane
and perform rotation and shifting in order to maximize SIFT feature
matches in overlapping areas between these projections. We then select
which textures to be used either with the tile caching method from
Section \ref{sec:mappingWithCaching}, or with the more specialized
method to be described in Section \ref{sec:seamMinimization}. This
choice depends on the availability of head-on images for a given
plane. We then finish by applying linear alpha blending to smooth out
any remaining seams.

\subsection{Image Rotation}
\label{sec:projectionAndRotation}
We begin with the projection of all images onto the plane. This is
done in the same way as the approaches in Section
\ref{sec:simpleTextureMapping}. Rather than work with all 6 DOF, which is computationally intensive, we choose to work only in two dimensions, performing 2D rotations and translations on the image projections, as errors in either lead to the greatest discontinuities.

We perform rotations using Hough transforms, which
detect the presence and orientation of linear features in our
images. Rather than match the orientation of such features in each
image, we simply apply rotations such that the strongest near-vertical
features are made completely vertical. This is effective for vertical
planes in indoor models, since they usually consist of parallel
vertical lines corresponding to doors, wall panels, rectangular
frames, etc. If features in the environment are not vertical, e.g. for
floors and ceilings, or are not parallel to each other, this step is
skipped.

\subsection{Image Shifting}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align overlapping images by searching for
corresponding points between all pairs of overlapping images. We use SIFT features for their high detection rate, and choose to use feature alignment rather than intensity-based alignment due to the differences in lighting as well as possible occlusion among our images, both of which feature alignment is less sensitive to \cite{lai1999robust, lowe1999object, mikolajczyk2005performance, szeliski2006image}.

SIFT matches determine $d^x$ and $d^y$ distances between each pair of
features for two images on the plane, though these distances may not
always be the same for different features.

Since indoor environments often contain repetitive features such as
floor tiles or doors, we need to ensure that SIFT-based distances are
reliable. In order to mitigate the effect of incorrect matches and
outliers, the RANSAC framework \cite{fischler1981random} is used for a
robust estimate of the optimal $d^x_{i,j}$ and $d^y_{i,j}$ distances
between two images $i$ and $j$. The RANSAC framework handles the
consensus-building machinery, and requires a fitting function and a
distance function. For this application, the fitting function simply
finds the average distance between matches in a pair of images. The
distance function for a pair of points is chosen to be the difference
between those points' SIFT match distance and the average distance
computed by the fitting function. We use a 10 pixel outlier threshold,
so that SIFT matches are labeled as outliers if their horizontal or
vertical distances are not within 10 pixels of the average distance
computed by the fitting function.

We now use the $d^x_{i,j}$ and $d^y_{i,j}$ distances between each pair
of images to refine their positions using least squares. Recall that
there are a total of $M^{2}$ possible pairs of images, though we only
generate distances between images that overlap at SIFT feature
points. Given these distances and the original image location
estimates, we can solve a least squares problem
($\textrm{min}_{\vec{\beta}} ||A \vec{\beta} - \vec{\gamma}||_2^2 $)
to estimate the location of the images on the plane. The
$M$-dimensional vector $\vec{\beta}$ represents the unknown $x$
location of each image on the plane for $1 \dots M$. The optimal $x$
and $y$ locations are obtained in the same way, so we only consider
the $x$ locations here:

@@

The $\vec{\beta}$ that minimizes $||A \vec{\beta} -
\vec{\gamma}||_2^2$ results in a set of image locations on the plane
that best honors all the SIFT-based distance measurements between
images. In practice there are often cases where there is a break in
the chain of images, meaning that no SIFT matches are found between
one segment of the plane and another. In this case we add rows to the
$A$ matrix and observations to the $\vec{\gamma}$ vector that contain
the original noisy $x$ and $y$ distance estimates generated by the
localization algorithm \cite{chen2010indoor, liu2010indoor}. Another
way to do this is to add rows for all neighboring pairs of images and
solve a weighted least squares problem where the SIFT distances are
given a higher weight i.e. 1, and the noisy distances generated by the
localization algorithm \cite{chen2010indoor, liu2010indoor} are given
a smaller weight i.e. 0.01.

After completing this same process for the $y$ dimension as well, and
making the resultant shifts, our images overlap and match each other
with far greater accuracy. Applying the tile caching method from
Section \ref{sec:mappingWithCaching} on these rotated and shifted
images results in the significant improvements shown in Figure
\ref{fig:compareAll}(c), as compared to Figure
\ref{fig:compareAll}(b).

@@

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=1in]{projectionWall.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=1.1in]{projectionCeiling.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=0.9in]{projectionWallCrop.pdf}}
  \caption{(a) Undistorted fisheye images for vertical planes are
    tilted, but their effective camera axis is more or less normal to
    the plane. (b) The effective camera axis for ceilings is at an
    angle with respect to the plane normal. (c) Wall images are
    cropped to be rectangular.}
  \label{fig:projectionAngles}
\end{figure}
